---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into multiple options: under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus. How many loans are in each group? You will need to use `case_when` to make this work and there's an example of this in the book chapter on mutate.
**A1** There are 9310 loans in the under 10k group, 15306 loans in the 10k to 50k range, 3968 loans in the 50k to 100k range, 4132 loans in the 100k go 500k range, 539 loans in the 500k to 1 million group, 356 loans in the 1m plus group and 97 loans did not fall into any of these categories. 

```{r}
west_virginia_ppp %>%
  mutate(
    loan_brackets = case_when(
        amount < 10000 ~ "under_10k",
        amount >= 10001 & amount <= 50000 ~ "10k_50k",
        amount >= 50001 & amount <= 100000 ~ "50k_100k",
        amount >= 100001 & amount <= 500000 ~ "100k_500k",
        amount >= 500001 & amount <= 1000000 ~ "500k_1m",
        amount > 1000001 & amount ~ "1m_plus",
        
      )
  )%>% 
  group_by(loan_brackets)%>%
    summarise(
        count_loans = n()
    )
  
```


**Q2** Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Harvest Small Business Finance, LLC has the most loans with undisbursed amounts greater than 0. 

```{r}
west_virginia_ppp %>%
  mutate(
    greater_than_0=if_else(
      undisbursed_amount>0, "yes", "no"
    )
  )%>%
  filter(greater_than_0 == "yes") %>%
  group_by(servicing_lender_name)%>%
  summarise(count_undisbersed_loans = n()
            ) %>%
  
  arrange(desc(count_undisbersed_loans))
```

**Q3** How would you describe the top 10 list of those lenders? What do they have in common or not? (You may want to look up some of the top ones.)
**A3** Some of the top lenders are smaller banks and financial institutions that are focused in providing residential and commercial mortgages. These financial corporations are very much rooted in community initiatives and supporting specific areas instead of a large consumer base. 

**Q4** Which month in 2021 had the most loans approved?
**A4** April had the most loans approved 

```{r}
west_virginia_ppp %>%
  mutate(
    month = floor_date(date_approved, "month")) %>%
  group_by(month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```
