---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("97677b7fa54c8bb69e6ed34e1b9841350e55ff82")
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
ACS5 <- load_variables(2015, "acs5", cache = FALSE)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
PPP_loan_applications <- read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first.
**A1**  Baltimore County is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates. 

```{r}
ACS2019 <- get_acs(geography = "county",
              variables = c(label="B01002_001"),
              state = "MD",
              year = 2019)
```

**Q2** Which Maryland counties have a median age of at least 50?
**A2** Talbot and Worchester Counties have a median age of at least 50. 

```{r}
Age50 <- get_acs(geography = "county",
              variables= c(label="B01002_001"),
              state = "MD",
              year = 2019) %>%
arrange(estimate)
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** Prince George's County has the highest percentage of 5-17 year olds that speak Spanish at home. 

```{r}
county_level <- get_acs(geography = "county",
              variables = c(label = "B16007_004"),
              state = "MD",
              year = 2019) %>%
rename(speakspan517=estimate)

total_pop <- get_acs(geography = "county",
              variables = c(label  = "B16007_002"),
              state = "MD",
              year = 2019) %>%
rename(total517=estimate)

total_pop %>%
  inner_join(county_level, by=c("GEOID", "NAME")) %>%
  select(-variable.x, -variable.y, -moe.x, -moe.y) %>%
  mutate(change = ((speakspan517)/total517)*100) %>%
  arrange(change)

```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4** Something that stands out to me abotu the data is the jump in numbers after Prince Georges County and Montgomery County. In Montgomery County, 20% of 5-17 year olds speak Spanish at home which is 36099 of 177565 and then in Caroline County, only 10 percent of 5-17 year olds speak Spanish at home which is 598 out of 5832. There are significantly less 5-17 year olds in tese counties. Garrett County has only 4078 5-17 year olds and	29 of them speak Spanish at home. This is a very low number of 5-17 year olds. 

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** 30097 is the zip code with the most removed PPP applications. 
```{r}

PPP_loan_applications %>%
  filter(race == "Black or African American") %>%
  filter(zip =="30097")

PPP_loan_applications <- PPP_loan_applications %>%
  mutate(zip5 = str_sub(zip, 1, 5))

top_zips <- PPP_loan_applications %>%
 group_by(zip5, race) %>%
 summarise(total=sum(amount)) %>%
 arrange(desc(total))




black_total <- get_acs(geography = "zip code tabulation area",
              variables = c(agetotal = "B01001B_001"),
              year = 2019)  
black_total <- black_total %>%
  select(-variable, -moe, -NAME) %>%
  rename(estimate1 = estimate)
totals <- get_acs(geography = "zip code tabulation area",
              variables = c(agetotal = "B01001_001"),
              year = 2019)  
totals <- totals %>%
  select(-variable, -moe, -NAME) %>%
  rename(estimate2 = estimate)
complete<- black_total %>%
  inner_join(totals)
complete <- complete %>%
  mutate( percent = (estimate1/estimate2)) %>%
  rename(zip5 = GEOID)
complete <- top_zips %>%
  left_join (complete)
 



```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed application in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6** Multiple of the top 10 zip codes with the most removed application in terms of Black population were in Chicago. There were some scattered ones including Utah, New York, Georgia and Texas. They are all from lower income areas. 

**Q7** What other columns from the PPP data could be useful in reporting on what happened to this removed loans?
**A7** Business type and lender are other columns from the PPP data that could be useful in reporting on what happened to the removed loans. 
