---
title: "Data Analysis Project"
author: "Kylie, Shreya, Michael"
date: "11/30/21"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are working with the [Washington Post police shooting database](https://github.com/washingtonpost/data-police-shootings) and the [Washington Post homicides database.](https://github.com/washingtonpost/data-homicides)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
#essential libraries
library(tidyverse)
library(janitor)
library(lubridate)
#additional libraries 
library(sf)
library(tigris)
library(tidycensus)

install.packages("usmap")
library(usmap)

install.packages("censusxy")
library(censusxy)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
police_shootings <-read_csv("data/data-police-shootings-master/fatal-police-shootings-data.csv")

census_api_key("97677b7fa54c8bb69e6ed34e1b9841350e55ff82")

```

#working with the data

``` {r}
police_shootings <- read_csv("data/data-police-shootings-master/fatal-police-shootings-data.csv")

data1 <- read_rds("data/geocoded_results_500.rds")
data2 <-read_rds("data/geocoded_results_1000.rds")
data3 <-read_rds("data/geocoded_results_1500.rds")
data4 <-read_rds("data/geocoded_results_2000.rds")
data5 <-read_rds("data/geocoded_results_2500.rds")
data6 <-read_rds("data/geocoded_results_3000.rds")
data7 <-read_rds("data/geocoded_results_3500.rds")
data8 <-read_rds("data/geocoded_results_4000.rds")
data9 <-read_rds("data/geocoded_results_4500.rds")
data10 <- read_rds("data/geocoded_results_5000.rds")
data11 <- read_rds("data/geocoded_results_5500.rds")
data12 <- read_rds("data/geocoded_results_6000.rds")
data13 <- read_rds("data/geocoded_results_6410.rds")

shootings_w_geoid <- rbind(data1,data2,data3,data4,data5,data6,data7,data8,data9,data10,data11,data12,data13)

census_data <- get_acs(geography = "tract",variables = c(population = "B01001_001", median.gross.rent = "B25064_001",median.household.income = "B19013_001", rent.burden = "B25071_001", white = "B03002_003", af.am = "B03002_004", hispanic = "B03002_012", am.ind = "B03002_005", asian = "B03002_006", nh.pi = "B03002_007", multiple = "B03002_009",other = "B03002_008"), state = c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY"), year = 2019)

#pivot

census_data <- pivot_wider(census_data, names_from = variable, names_sep = ".", values_from = c(estimate, moe)) 

census_data <- rename(census_data, census_tracts_geoid = GEOID)


unmatched <- anti_join(police_shootings, shootings_w_geoid, by=c("id"))

#join
shootings_w_acs <- shootings_w_geoid %>%
  left_join(census_data, by=c("census_tracts_geoid"))

new_shootings_w_acs <- read_rds("data/new_shootings_w_acs copy.rds")

new_shootings_w_acs <- new_shootings_w_acs %>%
  mutate(tract_majority_black = case_when((estimate.af.am/estimate.population > .5) ~ "y",
                                          (estimate.af.am/estimate.population <= .5) ~ "n")) %>%
  mutate(tract_majority_white = case_when((estimate.white/estimate.population > .5) ~ "y",
                                          (estimate.white/estimate.population <= .5) ~ "n")) %>%
  mutate(tract_majority_poor = case_when((estimate.poverty_level/estimate.population > .25) ~ "y",
                                         (estimate.poverty_level/estimate.population <= .25) ~ "n"))

##below poverty and Black shootings

below_pov_black <- new_shootings_w_acs %>%
  filter(
    tract_majority_black == "y",
    tract_majority_poor == "y"
  )
(395/72877) * 100


##.5420091% of shootings happen in tracts where a majority of the population is below the poverty line and a majority of the population is black.

tracts_below_black <- grouped_shootings %>%
  filter(
  tract_low_income == "y", 
  tract_majority_black == "y"
  )
(1447/72877) * 100

##1.985537% of all U.S. census tracts have a majority of the population that is below the poverty line AND a majority of the population that is black. 



##below poverty and white shootings

below_pov_white <- new_shootings_w_acs %>%
  filter(
    tract_majority_white == "y",
    tract_majority_poor == "y"
  )
(11/72877) * 100
  = 0.01509393% 

##0.01509393% of shootings happen in tracts where a majority of the population is white and a majority of the population is below the poverty line. 


grouped_shootings <- read_rds("data/grouped_shootings_w_acs.rds")






#deeper analysis
black_neighborhood_shootings <- new_shootings_w_acs %>%
  mutate(
    pct_black = (estimate.af.am/estimate.population) * 100)  %>%
    filter(pct_black > 50) %>%
  select(name, age, date, race, city, state, NAME, pct_black) %>%
  arrange(desc(pct_black))

(682 / 6409) * 100
## interesting --> 10.64129 percent of the total shootings were in majority black neighborhoods. 
## how many of those were black people? 
black_neighborhood_shootings %>%
  filter(race == "B") %>%
  summarize (
    count = n() 
  ) %>%
  as_tibble()
## 508 of those are black 
(508 / 682) * 100 
  = 74.4868 percent of the shootings in Black neighborhoods

## racial breakdown of the shootings
black_neighborhood_shootings %>%
  filter(race == "B") %>%
  summarize (
    count = n() 
  ) %>%
  as_tibble()


##white neighborhood breakdown of shootings by race

##white people shot in majority white neighborhoods
white_neighborhood_shootings_white <- new_shootings_w_acs %>%
  filter(race == "W", estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)

##Black people shot in majority white neighborhoods
white_neighborhood_shootings_black <- new_shootings_w_acs %>%
  filter(race == "B", estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)

## total people shot in majority white neighborhoods
majority_white_neighborhood_shootings <- new_shootings_w_acs %>%
  filter(estimate.white > estimate.af.am + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)

2573/4856 = 0.52986 
    = 52.986% of people shot in majority white neighborhoods are white. 

709/4856 = 0.1460049
    = 14.6% of people shot in majority white neighborhoods are black. 

## black people shot in majority black neighborhoods
black_neighborhood_shootings_black <- new_shootings_w_acs %>%
filter(race == "B", estimate.af.am > estimate.white + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
black_neighborhood_shootings_black
  ##663 people

## white people shot in majority black neighborhoods
black_neighborhood_shootings_white <- new_shootings_w_acs %>%
filter(race == "W", estimate.af.am > estimate.white + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
black_neighborhood_shootings_white
  ##144 people

## total people shot in majority black neighborhoods
majority_black_neighborhood_shootings <- new_shootings_w_acs %>%
filter(estimate.af.am > estimate.white + estimate.am.ind + estimate.asian + estimate.nh.pi + estimate.other)
majority_black_neighborhood_shootings
  ##1029 people
144/1029 = 0.1399417*100 
    = 13.4% of people shot in majority black neighborhoods are white. 
663/1029 = 0.6443149*100
    = 64.4% of people shot in majority black neighborhoods are black. 
```



``` {r}
##back to square 1 :P 

shooting_threat_level <- new_shootings_w_acs %>%
  filter(threat_level == "attack")

(4119/6409) *100

##64.269% of shootings happened when the threat level was attack. 


fleeing <- new_shootings_w_acs %>%
  filter(flee == "Not fleeing")

(3829/6409) *100

##59.74411 people were not fleeing when they were shot. 

car_fleeing <- new_shootings_w_acs %>%
  filter(flee == "Car")
(1012/6409) *100

##15.79029 people were fleeing by car.


```



### Income Analysis


``` {r}
## income breakdown !! 
    ## Pew Research Center defines low-income households as those earning below $48,500 
    ## Pew Research Center defines high-income households as those earning above $145,500
    
## general low-income breakdown first:
low_income_shootings <- new_shootings_w_acs %>%
  filter(estimate.median.household.income < 48500, 
         race == "B")
low_income_shootings %>% 
   summarize(
   count = n() 
  ) 
  ## 2924 shootings out of 6409 were in low-income neighborhoods 
(2924/6409) * 100
  = 45.6% of shootings were in low-income neighborhoods.
(865/6409) * 100
  = 13.49665 % of shootings happen in tracts where the majority of the population is below the poverty line and a majority of the population is black. 

## where are the specific places:
low_income_shootings %>%
  group_by(NAME) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))
  ##none of these seem to be significant

## white people shot in majority low-income neighborhoods
low_income_neighborhood_shootings_white <- new_shootings_w_acs %>%
filter(race == "W", estimate.median.household.income < 48500)
low_income_neighborhood_shootings_white
  ##1108 people

## black people shot in majority low-income neighborhoods
low_income_neighborhood_shootings_black <- new_shootings_w_acs %>%
filter(race == "B", estimate.median.household.income < 48500)
low_income_neighborhood_shootings_black
  ##865 people
1108/2924 = 0.378933*100 
    = 37.9% of people shot in majority low-income neighborhoods are white. 
865/2924 = 0.2958276*100
    = 29.6% of people shot in majority low-income neighborhoods are black. 
## general high-income breakdown
high_income_shootings <- new_shootings_w_acs %>%
  filter(estimate.median.household.income > 145500)
high_income_shootings %>% 
   summarize(
   count = n() 
  ) 
  ## 59 shootings out of 6409 were in high-income neighborhoods 
(59/6409) * 100
  = 0.92% of shootings were in high-income neighborhoods.
## where are the specific places:
high_income_shootings %>%
  group_by(NAME) %>%
   summarize(
   count = n() 
  ) %>%
  arrange(desc(count))
  ##none of these seem to be significant, except maybe top 3 in Texas?
## white people shot in majority high-income neighborhoods
high_income_neighborhood_shootings_white <- new_shootings_w_acs %>%
filter(race == "W", estimate.median.household.income > 145500)
high_income_neighborhood_shootings_white
  ##33 people
## black people shot in majority high-income neighborhoods
high_income_neighborhood_shootings_black <- new_shootings_w_acs %>%
filter(race == "B", estimate.median.household.income > 145500)
high_income_neighborhood_shootings_black
  ##13 people
33/59 = 0.378933*100 
    = 55.9% of people shot in majority high-income neighborhoods are white. 
13/59 = 0.2958276*100
    = 22.0% of people shot in majority high-income neighborhoods are black. 







### 1. What percentage of shootings happen in low_income tracts?
new_shootings_w_acs %>% 
  filter(tract_low_income == "y") %>%
   summarize(
    count = n()
  )
## 3465 shootings in low-income tracts, 6409 shootings in total
(3465 / 6409) * 100 
## 54.064 percent of shootings happen in low-income tracts

### 2. What percentage of all U.S. census tracts are low_income
grouped_shootings_w_acs %>% 
  group_by(tract_low_income) %>%
  filter(!is.na(tract_low_income)) %>%
  summarize(
    count = n()
  )
(49283	/ (49283 + 22749)) * 100
## 68.41 percent of tracts are low-income

### 3. How does the answer to number 1 compare to the answer in number 2?
## this might not be super significant since most of the U.S. is low-income tracts--could be indicative of our benchmark

# All low_income and majority black tracts

### 1. What percentage of shootings happen in tracts that are low_income AND majority black?
new_shootings_w_acs %>%
  filter(tract_low_income == "y" & tract_majority_black == "y") %>%
  summarize(count = n())
(101 / 6409) * 100
## 1.57 percent happen in tracts that are low-income and majority black

### 2. What percentage of all U.S. census tracts are low_income and majority black?
grouped_shootings_w_acs %>%
  filter(tract_low_income == "y" & tract_majority_black == "y") %>%
  summarize(count = n())
(1447 / 73056) * 100
## 1.98 percent of census tracts are low-income and majority black

### 3. How does the answer to number 1 compare to the answer in number 2?
## about the same. 

# All low_income and majority white tracts  

#1. What percentage of shootings happen in tracts that are low_income AND majority white?
new_shootings_w_acs %>%
  filter(tract_low_income == "y" & tract_majority_white == "y") %>%
  summarize(count = n())
  #2391 
2391/6409 = .3730691*100
    = 27.3% of shootings happen in tracts that are low_income AND majority white.

#2. What percentage of all U.S. census tracts are low_income and majority white?
grouped_shootings_w_acs %>%
  filter(tract_low_income == "y" & tract_majority_white == "y") %>%
  summarize(count = n())
  #37842 
37842/73056 = .5179862*100
    = 51.8% of all U.S. census tracts are low income and majority white.

#3. How does the answer to number 1 compare to the answer in number 2?
  





# Using poverty
## All majority poor tracts
#1. What percentage of shootings happen in tracts where the majority of the population is below the poverty line? 
new_shootings_w_acs %>%
  filter(tract_majority_poor == "y") %>%
  summarize(count = n())
  #90 
90/6409 = .01404275*100
    = 1.4% of shootings happen in tracts where the majority of people are below the poverty line.

#2. What percentage of all U.S. census tracts have a majority of the population below the poverty line?

grouped_shootings_w_acs %>%
  filter(tract_majority_poor == "y") %>%
  summarize(count = n())
  #745 
745/73056 = .01019766*100
    = 1.0% of all U.S. census tracts are low income and majority white.








## All majority poor and majority black tracts
1. What percentage of shootings happen in tracts where a majority of the population is below the poverty line AND a majority of the population is black?

below_pov_black <- new_shootings_w_acs %>%
  filter(
    tract_majority_black == "y",
    tract_majority_poor == "y"
  )
(395/72877) * 100

##54.20091% of shootings happen in tracts where a majority of the population is below the poverty line and a majority of the population is black.


2. What percentage of all U.S. census tracts have a majority of the population that is below the poverty line AND a majority of the population that is black?

tracts_below_black <- grouped_shootings %>%
  filter(
  tract_low_income == "y", 
  tract_majority_black == "y"
  )
(1447/72877) * 100

1.985537% of all U.S. census tracts have a majority of the population that is below the poverty line AND a majority of the population that is black. 


3. How does the answer to number 1 compare to the answer in number 2?

## All majority poor and majority white tracts  
1. What percentage of shootings happen in tracts where a majority of the population is below the poverty line AND a majority of the population is white?

below_pov_white <- new_shootings_w_acs %>%
  filter(
    tract_majority_white == "y",
    tract_majority_poor == "y"
  )
(11/72877) * 100
  = 0.01509393% 

0.01509393% of shootings happen in tracts where a majority of the population is white and a majority of the population is below the poverty line. 


2. What percentage of all U.S. census tracts have a majority of the population that is below the poverty line AND a majority of the population that is white?
  
3. How does the answer to number 1 compare to the answer in number 2?  
 


  
 
  
  
  
# State breakdown
# Do all of the queries above, but instead of asking the question nationally, ask it for each state (write one query to do it for all states, don't write a spearate query for each state)


```