---
title: "Data Analysis Project"
author: "Kylie, Shreya, Michael"
date: "11/14/21"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are working with the [Washington Post police shooting database](https://github.com/washingtonpost/data-police-shootings) and the [Washington Post homicides database.](https://github.com/washingtonpost/data-homicides)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
#essential libraries
library(tidyverse)
library(janitor)
library(lubridate)
#additional libraries 
library(sf)
library(tigris)
library(tidycensus)
install.packages("usmap")
library(usmap)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
police_shootings <- read_csv("data/data-police-shootings-master/fatal-police-shootings-data.csv")
homicides <- read_csv("data/data-homicides-master/homicide-data.csv")

census_api_key("97677b7fa54c8bb69e6ed34e1b9841350e55ff82")
```

## Working with the data 

Question 1: Is there a correlation between those fatally shot who were armed with a toy weapon and race?

``` {r}
## trying to make a map to generally understand where the shootings are
## getting some county data
counties <- counties() 
usa_counties <- counties 
## adding a geometry columns to police shootings
police_shootings_geom <- police_shootings %>%
  mutate(geometry = c(latitude, longitude))
## above did not work 
 
## mapping all the counties
## this is not really working rn but I am trying lol
ggplot() + 
  geom_sf(data=usa_counties) + 
  geom_sf(data=police_shootings) +
  theme_minimal() +
  geometry = geometry
```

Question 2: We know that police disproportionately kill Black men. Are these killings happening in Black neighborhoods or elsewhere? Are police shootings disproportionately happening in low-income areas?

``` {r}

```

Question 3: 
``` {r}

```

Question 4: 

``` {r}
```

Question 5: 

``` {r}

```
